---
title: "KMeans Clustering"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

http://www.business-science.io/business/2016/08/07/CustomerSegmentationPt1.html

Example, trying to cluster bike shop customers based on their bike buying preferences.Finds groups based on k-means euclidean distance. The closer they are to each other, the shorter the distance, and the more likely they are to be clumped together into a group.


```{r}
# Read Cannondale orders data --------------------------------------------------
library(xlsx) 
library(cluster)

customers <- read.xlsx("C:/Users/nrb75/Documents/KMeansEX/orderSimulatoR/data/bikeshops.xlsx", sheetIndex = 1)
products <- read.xlsx("C:/Users/nrb75/Documents/KMeansEX/orderSimulatoR/data/bikes.xlsx", sheetIndex = 1) 
orders <- read.xlsx("C:/Users/nrb75/Documents/KMeansEX/orderSimulatoR/data/orders.xlsx", sheetIndex = 1) 
```


Merge data together
```{r}
# Combine orders, customers, and products data frames --------------------------
library(dplyr)
orders.extended <- merge(orders, customers, by.x = "customer.id", by.y="bikeshop.id")
orders.extended <- merge(orders.extended, products, by.x = "product.id", by.y = "bike.id")

orders.extended <- orders.extended %>%
  mutate(price.extended = price * quantity) %>%
  select(order.date, order.id, order.line, bikeshop.name, model,
         quantity, price, price.extended, category1, category2, frame) %>%
  arrange(order.id, order.line)

knitr::kable(head(orders.extended)) # Preview the data
```


We hypothesize that bike shops purchase bike models based on features (mtn or road) and price (high or low). We will use bike model to cluster on, because the bike model contains this information i.e, price, what style, mtn, etc. We will then look at the preferences based on which bike models fall into certain groups. 
The unit of measure to cluster on = quantity of specific bike model purchased.

Manipulate dataframe for clustering, 
```{r}
# Group by model & model features, summarize by quantity purchased -------------
library(tidyr)  # Needed for spread function
customerTrends <- orders.extended %>%
        group_by(bikeshop.name, model, category1, category2, frame, price) %>%
        summarise(total.qty = sum(quantity)) %>%
        spread(bikeshop.name, total.qty)
customerTrends[is.na(customerTrends)] <- 0  # Remove NA's
```

We’ll segment the price into high/low by median price. Selecting g = 2 divides the unit prices into two halves using the median as the split point.
```{r}
# Convert price to binary high/low category ------------------------------------
library(Hmisc)  # Needed for cut2 function
customerTrends$price <- cut2(customerTrends$price, g=2)   

```

normalize the # of bikes by the total # purchased at each shop. This controls for differences in stores (one can buy 100 bikes a year and another 10,000)

```{r}
# Convert customer purchase quantity to percentage of total quantity -----------
customerTrends.mat <- as.matrix(customerTrends[,-(1:5)])  # Drop first five columns
customerTrends.mat <- prop.table(customerTrends.mat, margin = 2)  # column-wise pct
customerTrends <- bind_cols(customerTrends[,1:5], as.data.frame(customerTrends.mat))

```


## K-means clustering
'k' number of groups to cluster customers.

```{r}
kmeansDat <- customerTrends[,-(1:5)]  # Extract only customer columns
kmeansDat.t <- t(kmeansDat)  # Get customers in rows and products in columns

# Setup for k-means loop 
km.out <- list()
sil.out <- list()
x <- vector()
y <- vector()
minClust <- 4      # Hypothesized minimum number of segments
maxClust <- 8      # Hypothesized maximum number of segments

# Compute k-means clustering over various clusters, k, from minClust to maxClust
for (centr in minClust:maxClust) {
        i <- centr-(minClust-1) # relevels start as 1, and increases with centr
        set.seed(11) # For reproducibility
        km.out[i] <- list(kmeans(kmeansDat.t, centers = centr, nstart = 50))
        sil.out[i] <- list(silhouette(km.out[[i]][[1]], dist(kmeansDat.t)))
        # Used for plotting silhouette average widths
        x[i] = centr  # value of k
        y[i] = summary(sil.out[[i]])[[4]]  # Silhouette average width
}
```

Next, we plot the silhouette average widths for the choice of clusters. The best cluster is the one with the largest silhouette average width, which turns out to be 5 clusters.


```{r}
# Plot silhouette results to find best number of clusters; closer to 1 is better
library(ggplot2)
ggplot(data = data.frame(x, y), aes(x, y)) + 
  geom_point(size=3) + 
  geom_line() +
  xlab("Number of Cluster Centers") +
  ylab("Silhouette Average Width") +
  ggtitle("Silhouette Average Width as Cluster Center Varies")
```

4 Clusters is best, b/c it has highest silhouette value

Now pull out customers in each cluster
```{r}
# Get customer names that are in each segment ----------------------------------

# Get attributes of optimal k-means output
maxSilRow <- which.max(y)          # Row number of max silhouette value
optimalClusters <- x[maxSilRow]    # Number of clusters
km.out.best <- km.out[[maxSilRow]] # k-means output of best cluster

# Create list of customer names for each cluster
clusterNames <- list()
clusterList <- list()
for (clustr in 1:optimalClusters) {
  clusterNames[clustr] <- paste0("X", clustr)
  clusterList[clustr] <- list(
    names(
        km.out.best$cluster[km.out.best$cluster == clustr]
        )
    )
}
names(clusterList) <- clusterNames

print(clusterList)
```


Inspect what makes each group different
```{r}
# Combine cluster centroids with bike models for feature inspection ------------
custSegmentCntrs <- t(km.out.best$centers)  # Get centroids for groups
colnames(custSegmentCntrs) <- make.names(colnames(custSegmentCntrs))
customerTrends.clustered <- cbind(customerTrends[,1:5], as.data.frame(custSegmentCntrs))

```

